---
title: "Homework#4"
author: "Hassan Fayyaz"
date: "3/13/2022"
output: pdf_document
---

'Important Note' | 'Homework Starts at line 358' 

#Cleaning the environment
```{r}
rm(list = ls())
```

#######################

# Loading Packages

#######################

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(rmarkdown)
library(tidyverse)
library(readr)
library(readxl)
```

```{r}
setwd("C:/Users/hassa/Desktop/Dataset")  

flower <- readxl::read_xls("flower.xls")

ggplot()


ggplot(data = flower,
       mapping = aes(x= weight, y= shootarea))
```

```{r}
#we need to introduce a geometry or layer that tells us how to display that data

ggplot(flower,
       aes(x = weight, y= shootarea)) + geom_point()
```

```{r}
# other options
ggplot(flower,
       aes(x= weight, y=shootarea)) + geom_line()

ggplot(flower,
       aes(x= weight, y = shootarea)) +
  geom_point()+
  ggtitle("Flower Weight vs Flower Shoot Area")+
  xlab("Weight")+
  ylab("Shoot Area")+
  theme_minimal()+
  #geom_smooth(method="lm", se = FALSE)+
  geom_smooth(se=FALSE)
```

```{r}
#add in additional variables
ggplot(flower,
       aes(x= weight, y= shootarea, color=nitrogen))+
  geom_point()+
  ggtitle("Flower Weight vs Flower Shoot Area")+
  xlab("Weight")+
  ylab("Shoot Area")+
  theme_minimal()+
  geom_smooth(aes(color = nitrogen), 
              method="lm", 
              se=FALSE)
```

```{r}
#4th
ggplot(flower,
       aes(x= weight, y= shootarea, color=nitrogen, shape = treat))+
  geom_point()+
  ggtitle("Flower Weight vs Flower Shoot Area")+
  xlab("Weight")+
  ylab("Shoot Area")+
  theme_minimal()
```

```{r}
#split into grids

ggplot(aes(x= weight, 
           y= shootarea, 
           color=nitrogen), data = flower)+
  geom_point()+
  geom_smooth(method="lm", se=FALSE) + 
  facet_wrap(~treat)
```

```{r}
#split into multiple grids for more than 1 variable
ggplot(aes(x= weight, 
           y= shootarea, 
           color=nitrogen), data = flower)+
  geom_point()+
  geom_smooth(method="lm", se=FALSE) + 
  facet_wrap(~treat + block)
```

```{r}
#similar to wrap we can do a grid

ggplot(aes(x= weight, 
           y= shootarea, 
           color=nitrogen), data = flower)+
  geom_point()+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(~treat + block)
```

```{r}
#reverse the facets

ggplot(aes(x= weight, 
           y= shootarea, 
           color=nitrogen), data = flower)+
  geom_point()+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(~block + treat)
```

```{r}
#Final format
ggplot(aes(x= weight,
           y=shootarea,
           color=nitrogen),
       data=flower)+
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  facet_grid(block ~ treat) +
  xlab("Weight of Flower (g)") + 
  ylab("Area of Shoot (cm^2)")
```

```{r}
#Fiddle with the Title formatting using bquote
ggplot(aes(x= weight,
           y=shootarea,
           color=nitrogen),
       data=flower)+
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(block ~ treat) +
  xlab("Weight of Flower (g)") + 
  ylab(bquote("Area of Shoot" ~ (cm^2)))+
  theme_minimal()
```

```{r}
## Improve the Labels
ggplot(aes(x= weight,
           y=shootarea,
           color=nitrogen),
       data=flower)+
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(block ~ treat) +
  xlab("Weight of Flower (g)") + 
  ylab(bquote("Area of Shoot" ~ (cm^2)))+
  theme_minimal()+
  labs(shape = "Nitrogen Concentration",
       color = "Nitrogen Concentration",
       caption = "Regression Assumptions are Unvalidated")
```

```{r}
# save the graphics and then updated

plot1 <- ggplot(aes(x= weight,
                    y=shootarea,
                    color=nitrogen),
                data=flower)

plot2 <- plot1 +
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(block ~ treat) +
  xlab("Weight of Flower (g)") + 
  ylab(bquote("Area of Shoot" ~ (cm^2)))+
  theme_minimal()+
  labs(shape = "Nitrogen Concentration",
       color = "Nitrogen Concentration",
       caption = "Regression Assumptions are Unvalidated")
```

```{r}
#adding a title
plot2 <- plot2 +
  labs(title = "Area of Shoot vs Weight of Flower")

plot2 <- plot2 +
  geom_hline(aes(yintercept = mean(shootarea)),
             size = .5,
             color = "black")

plot2 <- plot2 +
  geom_hline(aes(yintercept = mean(shootarea)),
             size = .5,
             color = "black",
             linetype = 3)


ggplot(aes(x= weight,
           y=shootarea,
           color=nitrogen),
       data=flower)+
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  geom_smooth(method="lm", se=FALSE) + 
  xlab("Weight of Flower (g)") + 
  ylab(bquote("Area of Shoot" ~ (cm^2)))+
  theme_minimal()+
  xlim(c(10, 15))

```

```{r}
# display numeric data as a color gradient
ggplot(aes(x = weight,
           y= shootarea), data = flower) +
  geom_point(aes(color = height,
                 shape = nitrogen))+
  scale_shape_manual(values = c(15,17,19))
```

```{r}
#add a different color gradient
ggplot(aes(x = weight,
           y= shootarea), data = flower) +
  geom_point(aes(color = height,
                 shape = nitrogen))+
  scale_shape_manual(values = c(15,17,19))+
  scale_color_gradient(low= "pink", high = "black")
```

```{r}
### change default grid values for x and y axis

ggplot(aes(x= weight,
                    y=shootarea,
                    color=nitrogen),
                data=flower)+
  geom_point(aes(shape = nitrogen),
             size = 2,
             alpha = 1)+
  geom_smooth(method="lm", se=FALSE) + 
  facet_grid(block ~ treat) +
  xlab("Weight of Flower (g)") + 
  ylab(bquote("Area of Shoot" ~ (cm^2)))+
  labs(shape = "Nitrogen Concentration",
       color = "Nitrogen Concentration",
       caption = "Regression Assumptions are Unvalidated")+
  scale_x_continuous(breaks = seq(from = 5, to = 25, by=5))+
  theme_minimal()
```

```{r}
### Density

ggplot(flower, 
       aes(x= flowers, 
           fill = nitrogen))+
  geom_density()+
  labs(y = "Density",
       x="Number of Flowers", 
       fill = "Nitrogen \n Concentration")+
  scale_fill_manual(labels = c("High",
                               "Medium",
                               "Low"),
                    values = c("dodgerblue",
                               "forestgreen",
                               "darkred"))+
  theme_minimal()
```

```{r}
#Histogram

ggplot(flower, 
       aes(x= flowers, 
           fill = nitrogen))+
  geom_histogram(position = "dodge")+
  labs(y = "Count",
       x="Number of Flowers", 
       fill = "Nitrogen \n Concentration")+
  scale_fill_manual(labels = c("High",
                               "Medium",
                               "Low"),
                    values = c("dodgerblue",
                               "forestgreen",
                               "darkred"))+
  theme_minimal()


ggplot(flower, 
       aes(x= flowers, 
           fill = nitrogen))+
  geom_histogram(position = "dodge")+
  facet_grid(~treat ~ block)+
  labs(y = "Count",
       x="Number of Flowers", 
       fill = "Nitrogen \n Concentration")+
  scale_fill_manual(labels = c("High",
                               "Medium",
                               "Low"),
                    values = c("dodgerblue",
                               "forestgreen",
                               "darkred"))+
  theme_minimal()

```

```{r}
#Frequency Polygon
ggplot(flower,
       aes(x= flowers,
           color = factor(nitrogen)))+
  geom_freqpoly(size = 1, bins= 20)+
  scale_color_manual(labels = c("Low", "Medium", "High"),
                     values = c("dodgerblue", "forestgreen", "darkred"))+
  theme_minimal()

```

```{r}
#Box plot

ggplot(flower,
       aes(y = flowers,
           x=nitrogen,
           fill = nitrogen))+
  geom_violin(draw_quantiles   = c(.25, .5, .75))+
  geom_jitter(aes(y= flowers,
                   x= nitrogen),
               color="black",
              height = 0,
              width=.1, alpha=.5)
```


```{r}
### heatmap

ggplot(aes(x = leafarea,
           y= shootarea),
       data = flower) +
  geom_bin2d()
```

```{r}
### contour mapping
ggplot(aes(x= leafarea,
           y= shootarea), data = flower) +
  geom_density2d()
```

```{r}
library(GGally)

flower$block <- factor(flower$block)

ggpairs(flower,
        aes(color = nitrogen))
```

```{r}
### time series
ggplot(economics,
       aes(x= date)) + 
  geom_line(aes(y= psavert), color = "dodgerblue")+
  geom_line(aes(y= uempmed), color="darkred")
```


#################################################################################

                                'Homework#4'

#################################################################################
'Instructions:

Find a dataset interesting to you that includes several numeric and catagorical 
variables build a numbers (at least 10) graphs from your data like the ones we 
did today. Make sure to add clean descriptive titles, axis labels and labels'
#################################################################################
                    'Getting Ready To Plot The Graphs'

```{r}
#Cleaning the environment:
rm(list = ls())
```

```{r}
#Changing the directory:
library(readxl)
whaledata <- read_excel("Dataset/whaledata.xls")
View(whaledata)
```

```{r}
#Installing Packages:
install.packages("tidyverse", dependencies = TRUE)
library(tidyverse)
```

```{r}
install.packages("ggplot2", dependencies = TRUE)
library(ggplot2)
```


```{r}
library(ggExtra)
```

#################################################################################
'GRAPH#1: Using SJPlots Function | Water Noise Based On The Number Of Whales'
#################################################################################
```{r}
#TABLE: (Water Noise Based On The Number Of Whales)
sjPlot::tab_xtab(var.row=whaledata$water.noise, 
                 var.col =whaledata$number.whales,
                 title = "Water Noise Based On The Number Of Whales",
                 show.row.prc = TRUE)

#GRAPH: (Water Noise Based On The Number Of Whales)
sjPlot::plot_xtab(whaledata$water.noise, whaledata$number.whales,
                  margin = "row",
                  bar.pos = "stack", coord.flip = TRUE)
```
#################################################################################
'Graph#2: Grouped Barplot | Water Noise Based On The Number Of Whales' 
#################################################################################
```{r}
#Water Noise Based On The Number Of Whales
counts <- table(whaledata$water.noise, whaledata$number.whales)
barplot(counts, main="Water Noise Based On The Number Of Whales",
        xlab="Number Of Whales", ylab="Noise", col=c("red","green","darkblue"),
        legend = rownames(counts), beside=TRUE)
```
#################################################################################
'Graph#3: Stacked Bar Plot | Water Noise Based On The Number Of Whales '
#################################################################################
```{r}
counts <- table(whaledata$water.noise, whaledata$number.whales)
barplot(counts, main="Water Noise Based On The Number Of Whales",
        xlab="Number Of Whales", ylab="Noise", col=c("red","green","darkblue"),
        legend = rownames(counts)) 
```
#################################################################################
'Graph#4: GGPLOT | Number of Whales vs Water Noise'
#################################################################################
```{r}
library(ggplot2)
ggplot(whaledata,
       aes(x= water.noise, y= number.whales))+
  geom_point()+
  ggtitle("Number of Whales vs Water Noise")+
  xlab("Month")+
  ylab("Number of Whales")+
  theme_minimal()
```
#################################################################################
'Graph#5: Violin GGPLOT | Water noise based on the depth'
#################################################################################
```{r}
ggplot(whaledata,
       aes(y = water.noise,
           x=depth))+
  geom_violin(draw_quantiles   = c(.25, .5, .75))+
  geom_jitter(aes(y= water.noise,
                  x= depth),
              color="Purple",
              height = 0,
              width=0.5, alpha=100)
```
#################################################################################
'Graph#6: GGPLOT For Time Spent At The Station Every Month'
#################################################################################
```{r}
ggplot(aes(x = month,
           y= time.at.station),
       data = whaledata) +
  geom_bin2d()
```
#################################################################################
'Graph#7: Multiple ggplot2 Charts On A Single Page '
#################################################################################
```{r}
# Make 3 simple graphics:
g1 <- ggplot(whaledata, aes(x=month)) + geom_density(fill="slateblue")
g2 <- ggplot(whaledata, aes(x=number.whales, y=month, color=water.noise)) + geom_point(size=5) + theme(legend.position="none")
g3 <- ggplot(whaledata, aes(x=factor(number.whales), y=month, fill=water.noise)) + geom_boxplot() + theme(legend.position="none")
g4 <- ggplot(whaledata , aes(x=factor(number.whales), fill=factor(water.noise))) +  geom_bar()
```

```{r}
# Plots
library(gridExtra)
grid.arrange(g2, arrangeGrob(g3, g4, ncol=2), nrow = 2)

grid.arrange(g1, g2, g3, nrow = 3)

grid.arrange(g2, arrangeGrob(g3, g4, ncol=2), nrow = 1)

grid.arrange(g2, arrangeGrob(g3, g4, nrow=2), nrow = 1)
```

#################################################################################
'Graph#8: A boxplot with jitter | Water noise based on the number of whales 
and depth of the ocean'
#################################################################################
```{r}
library(tidyverse)
library(hrbrthemes)
library(viridis)
```

```{r}
whaledata %>%
  ggplot( aes(x=depth, y=water.noise, fill=number.whales)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="Red", size=1.9, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Water noise based on the number of whales 
and depth of the ocean") +
  xlab("")
```
#################################################################################
'Graph#9: GGPLOT to understand the water noise based on the latitude & longitude'
##################################################################################
```{r}
ggplot(whaledata, mapping = aes(x = longitude, y = latitude)) +
  geom_point(alpha = 0.9, size=3, aes(color = water.noise))
```
#################################################################################
'Graph#10: Scatterplot matrix with ggpairs() '
#################################################################################
```{r}
library(GGally)
whaledata <- data.frame( var1 = 1:100 + rnorm(100,sd=20), v2 = 1:100 + rnorm(100,sd=27), v3 = rep(1, 100) + rnorm(100, sd = 1)) 
whaledata$v4 = whaledata$var1 ** 2 
whaledata$v5 = -(whaledata$var1 ** 2) 

ggpairs(whaledata, title="correlogram with ggpairs()") 
```
#################################################################################


