---
title: "Empirical Research HW#3"
author: "Hassan Fayyaz"
date: "2/22/2022"
output: pdf_document
---


```{r}
#Cleaning the environment
rm(list = ls())
```

#######################

# Loading Packages

#######################

```{r}
library(dplyr)
```

```{r}
library(ggplot2)
```

```{r}
library(tidyr)
```

```{r}
library(lubridate)
```

```{r}
library(rmarkdown)
```


#################################################################################

1) Find the natural log, log to the base 10, square root and the natural 
anti-log of 12.43

#################################################################################

```{r}
#natural log
log(12.43)

#log to the base of 10
log10(12.43)

# square root
sqrt(12.43) 

# natural anti-log
exp(12.43)
```


#################################################################################

2) Determine the area of a circle with a diameter of 50 inches and assign the 
result to a variable called area_circle. (Hint, use the pi function)

#################################################################################

```{r}
#Formula:  Area of a circle = pi × r2
#Radius: Diameter/2 | Which is 50/2 = 25

area_circle <- pi * (25)^2
print(area_circle)
```

#################################################################################

3) Use the concatenate function c() to create a vector called weight containing 
the weight (in lb) of 10 children: 120, 117, 139, 129, 94, 86, 106, 107, 126.  
Do the same for height in inches; 47, 40, 41, 35, 60, 42, 50, 49, 38, 42.  
Then calculate the bmi (weight/ the square of the height) of the kids.

#################################################################################

```{r}
weight <- c(140, 120, 117, 139, 129, 94, 86, 106, 107, 126)
height <- c(47, 40, 41, 35, 60, 42, 50, 49, 38, 42)

bmi <- (weight/(height)^2) * 703 
print(bmi)
```

#################################################################################

4) Create a new variable defined as name;, “John”, “Megan”, “Ryan”, “Isabela”, 
“Matt”, “Lindsay”, “Tony”, “Mary Jane”, “Chase” and “Delaney”.  Create a 
dateframe of name, height, weight, bmi and call it alldata.

#################################################################################

```{r}
name <- c("John", "Megan", "Ryan", "Isabela", "Matt", "Lindsay", "Tony", 
          "Mary Jane", "Chase", "Delaney")

alldata <- data.frame(c("name", "height", "weight", "bmi"))
```

#################################################################################

5) Sort the dataframe from tallest to shortest. Then create a new variable 
height_rank, from the tallest (1) to the shortest (2).  Repeat the same for 
weight, bmi, and name (alphabetically).  For bmi, order it from lowest to 
highest. (Hint, use the seq() function).

#################################################################################

```{r}
#Sorting the dataframe from tallest to shortest
alldata <- alldata[order(alldata$height, decreasing=TRUE),]

#Now creating new variables
height_rank <- rev(sort(height))
print(height_rank)

weight_rank <- rev(sort(weight))
print(weight_rank) 

bmi_rank <- (sort(bmi))
print(bmi_rank)

name_rank <- rev(sort(name))
print(name_rank)
```

#################################################################################

6)  Calculate the mean and standard deviation of the height and weight 
variables above and store them as height_mean, height_sd, weight_mean, 
weight_sd.  Using the rnorm function, create two new variables, height2 and 
weight2 by simulating 10,000  random normal variables with the same mean and 
standard deviation of the original data sets.

#################################################################################

```{r}
#Mean & Standard Deviation of Height

height_mean <- mean(height)
print(height_mean)

height_sd <- sd(height)
print(height_sd)
```

```{r}
#Mean & Standard Deviation of Weight

weight_mean <- mean(weight)
print(weight_mean)

weight_sd <- sd(weight) 
print(weight_sd)
```

```{r}
#Two new variables

height2 <- rnorm(10000, mean = 44.4, sd = 7.260242)
print(height2)

weight2 <- rnorm(10000, mean = 116.4, sd = 18.11813)
print(weight2)
```

#################################################################################

7) Create 10,000 fake names, name_1, name_2 ... name_10000. (Hint: use the 
paste() function and the seq() function together).  Call this name2.

#################################################################################

```{r}
names2 <- paste0("name_", seq(1:10000))
names2 <- paste("name_", seq(1:10000), sep="")
names2 <- paste("name", seq(1:10000), sep="_")
names2 < paste0("n", seq(1:10000), sep="ame_")
names2 <- paste0("names_", c(1:10000))
print(names2)
```

#################################################################################

8) Combine name2, height2, weight2 into a data.frame called alldata2.  
Calculate the bmi of the fake students and create a column from it.  
Then rename the all of the column names so that they match the first 
data.frame.

#################################################################################

```{r}
#Combining into a data.frame
alldata2 <- data.frame(name2, height2, weight2)
print(alldata2)
```

```{r}
#Now calculating the BMI of fake students
alldata2$bmi <- cbind(names2, alldata2$weight/alldata2$height^2 * 703)

#Adding heading for the table
colnames(alldata2$bmi) <- c("name2", "bmi2")

#printing the table
print(alldata2$bmi)
```

#################################################################################

9) Compare the BMI of the new dataset to the old data set.  Notice anything 
interesting?

#################################################################################

Results: 

After comparison, we have observed that the participants in experiment #1 
are more likely to be healthy as the majority of the participants have a BMI 
closer to 24.5, which is considered normal BMI. Whereas participants in 
experiment #2 are more likely to be obese.

#################################################################################



